---
title: "Heatmaps: version 1"
output: github_document
---

**Goal:** To create some cool heatmaps of the odds ratios of neighborhood transitions

Using the following packages:

```{r}
rm(list=ls())

library(corrplot)
library(reshape2)
library(ggplot2)
library(plyr)
library(zoo)
library(RColorBrewer)
```


### Steps to creating a heat map for 1990-2000
\\

**Step One:** Use the data from the previous analysis to create the neighborhood transition matrix.

```{r}
m <- matrix(c(0.1234, 0, 0, 0.0072, 0.0134, 0.0332, 0, 0, 0, 0.0022, 0.0048, 0.004, 0.0014, 0, 0.0468, 0, 0.001, 0, 0, 0.0042, 0.0004, 0, 0, 0.0006, 0, 0.0024, 0, 0, 0.0466, 0, 0.0006, 0, 0.0004, 0, 0.0008, 0, 0, 0.0002, 0, 0.0086, 0.0028, 0, 0.0502, 0.0004, 0.003, 0.0008, 0.0002, 0, 0.0064, 0.0122, 0.0002, 0.0022, 0.0026, 0, 0.0046, 0, 0.0466, 0.0022, 0.0004, 0.0002, 0.0004, 0.0044, 0, 0.0112, 0.0024, 0.0196, 0, 0, 0.0034, 0.0032, 0.101, 0, 0, 0, 0.001, 0.0234, 0.0164, 0.0086, 0, 0.0008, 0.0012, 0.0002, 0, 0, 0.0178, 0.001, 0, 0.0006, 0.0002, 0, 0.0004, 0, 0.0004, 0.0002, 0, 0, 0, 0.0044, 0.0062, 0.0008, 0, 0, 0, 0.0006, 0, 0, 0.004, 0, 0, 0, 0.0002, 0.0002, 0.0024, 0, 0, 0, 0, 0.0002, 0.0002, 0.0004, 0.0004, 0.0026, 0, 0.0038, 0.0004, 0, 0.0122, 0.0002, 0.0006, 0.0022, 0.0008, 0.0036, 0, 0.0116, 0, 0.0054, 0.0008, 0.0026, 0, 0.0026, 0.0584, 0.0006, 0.0246, 0.0014, 0, 0.0026, 0, 0.0144, 0.0078, 0.0004, 0.001, 0.0046, 0.0026, 0.0006, 0.059, 0.0186, 0, 0.005, 0.0004, 0.0002, 0.0008, 0.0008, 0.0036, 0.016, 0.0008, 0.005, 0.0022, 0.0026, 0.0406), nrow=13, ncol=13)

rownames(m) <- c("W", "B", "L", "WB", "WL", "WA", "BL", "BA", "LA", "WBL", "WBA", "WLA", "WBLA")
colnames(m) <- c("W", "B", "L", "WB", "WL", "WA", "BL", "BA", "LA", "WBL", "WBA", "WLA", "WBLA")

#reorder matrix
col.order <- c("W", "WB", "WBL", "WBA", "WBLA", "WL", "WLA", "WA", "B", "BL", "BA", "L", "LA")
row.order <- c(c("W", "WB", "WBL", "WBA", "WBLA", "WL", "WLA", "WA", "B", "BL", "BA", "L", "LA"))
m <- m[row.order,col.order]
```
\\

**Step Two:** Choose a pretty color scheme (Use color brewer!)

```{r}
## set color representation for specific values of the data distribution
quantile_range <- quantile(m, probs = seq(0, 1, 0.20))

## What color scheme do you want to use?
## Choose using the color brewer package
display.brewer.all(colorblindFriendly = T) #get colorblindfriendly colors
display.brewer.pal(n = 9, name = "PuRd")# Visualize a single RColorBrewer palette
brewer.pal(n = 9, name = "PuRd") # Return the hexadecimal color code of the palette

color_palette <- colorRampPalette(c("#F1EEF6", "#D7B5D8", "#DF65B0", "#DD1C77", "#67001F"))(length(quantile_range) - 1)
```
\\

**Step Three:** Figure out your aesthetic (aka make your plot readable and nice to look at)

```{r}
## prepare label text (use two adjacent values for range text)
label_text <- rollapply(round(quantile_range[-1], 5), width = 2, by = 1, FUN = function(i) paste(i, collapse = " : "))

## discretize matrix; this is the most important step, where for each value we find category of predefined ranges (modify probs argument of quantile to detail the colors)
mod_m <- matrix(findInterval(m, quantile_range, all.inside = TRUE), nrow = nrow(m))
rownames(mod_m) <- c("W", "WB", "WBL", "WBA", "WBLA", "WL", "WLA", "WA", "B", "BL", "BA", "L", "LA")
colnames(mod_m) <- c("W", "WB", "WBL", "WBA", "WBLA", "WL", "WLA", "WA", "B", "BL", "BA", "L", "LA")

## remove background and axis from plot
theme_change <- theme(
  plot.background = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_blank(),
  panel.background = element_blank(),
  panel.border = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.title.x = element_blank(),
  axis.title.y = element_blank()
)
```
\\

**Step Four:** Create the heat map using `ggplot2`

```{r}
# Heatmap 1
#png("heatmap1.png", width = 800, height = 400)
base_size <- 16 #text size
ggplot(melt(mod_m), 
                aes(x=Var1, y=Var2, fill=factor(value))) + 
  geom_tile(colour="white") + 
  scale_fill_manual(values = c("#F1EEF6", "#D7B5D8", "#DF65B0", "#980043"), 
                    name = "Probability", 
                    labels = c("0 : 0.0002", "0.0002 : 0.001", "0.001 : 0.0046", "0.0046 : 0.1234")) +
  theme_grey(base_size = base_size) + 
  labs(x = "2000 Neighborhood Type", y = "1990 Neighborhood Type") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.ticks=element_blank(),
        axis.title.x = element_text(size = 18, face = "bold"),
        axis.title.y = element_text(size = 18, face = "bold"))
#dev.off()
```


### Now repeat for for 2000-2010!

Using this matrix:

```{r}
m2 = matrix(c(0.0958,0, 0,0.008, 0.0148, 0.0222, 0, 0, 0, 0.003, 0.006, 0.0046, 0.0022, 0, 0.048, 0, 0.0014, 0, 0, 0.0048, 0.0004, 0, 0, 0.001, 0, 0.0038, 0, 0, 0.0562, 0, 0.0014, 0, 0.0008, 0, 0.001, 0, 0, 0.0006, 0, 0.0062, 0.0016, 0, 0.0462, 0.0004, 0.0018, 0.0006, 0.0002, 0, 0.0056, 0.0096, 0.0002, 0.0022, 0.003, 0, 0.005, 0.0002, 0.0518, 0.0018, 0.0004, 0.0002, 0.0004, 0.0054, 0.0002, 0.0106, 0.0032, 0.0154, 0, 0, 0.0032, 0.0032, 0.0892, 0, 0, 0, 0.0012, 0.0196, 0.0126, 0.0088, 0, 0.0012, 0.0024, 0.0002, 0.0002, 0, 0.028, 0.0012, 0.0002, 0.0014, 0.0004, 0.0002, 0.001, 0, 0.0004, 0.0006, 0, 0, 0, 0.0072, 0.016, 0.0014, 0.0002, 0.0002, 0, 0.0018, 0, 0, 0.0048, 0, 0.0002, 0, 0.0002, 0.0002, 0.0044, 0, 0, 0.0002, 0.0002, 0.0004, 0.0002, 0.0006, 0.0008, 0.0042, 0.0002, 0.0048, 0.0006, 0, 0.0194, 0.0004, 0.0012, 0.0042, 0.0008, 0.0022, 0, 0.0094, 0, 0.0034, 0.0006, 0.0018, 0, 0.0028, 0.0586, 0.0006, 0.0222, 0.0014, 0, 0.0024, 0, 0.0128, 0.0048, 0.0004, 0.0008, 0.003, 0.003, 0.0006, 0.0478, 0.018, 0.0002, 0.0054, 0.0008, 0.0006, 0.0014, 0.001, 0.0054, 0.0164, 0.001, 0.0092, 0.0042, 0.0044, 0.0538), nrow=13, ncol=13)

rownames(m2) <- c("W", "B", "L", "WB", "WL", "WA", "BL", "BA", "LA", "WBL", "WBA", "WLA", "WBLA")
colnames(m2) <- c("W", "B", "L", "WB", "WL", "WA", "BL", "BA", "LA", "WBL", "WBA", "WLA", "WBLA")
```

```{r, include=FALSE}

## set color representation for specific values of the data distribution
quantile_range2 <- quantile(m2, probs = seq(0, 1, 0.20))

#set color palette
color_palette2 <- colorRampPalette(c("#F1EEF6", "#D7B5D8", "#DF65B0", "#DD1C77", "#67001F"))(length(quantile_range2) - 1)

## prepare label text (use two adjacent values for range text)
label_text2 <- rollapply(round(quantile_range2[-1], 5), width = 2, by = 1, FUN = function(i) paste(i, collapse = " : "))

## discretize matrix; this is the most important step, where for each value we find category of predefined ranges (modify probs argument of quantile to detail the colors)
mod_m2 <- matrix(findInterval(m2, quantile_range2, all.inside = TRUE), nrow = nrow(m))
rownames(mod_m2) <- c("W", "B", "L", "WB", "WL", "WA", "BL", "BA", "LA", "WBL", "WBA", "WLA", "WBLA")
colnames(mod_m2) <- c("W", "B", "L", "WB", "WL", "WA", "BL", "BA", "LA", "WBL", "WBA", "WLA", "WBLA")

## remove background and axis from plot
theme_change <- theme(
  plot.background = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_blank(),
  panel.background = element_blank(),
  panel.border = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.title.x = element_blank(),
  axis.title.y = element_blank()
)
```

```{r}
# Heatmap 2
#png("heatmap2.png", width = 800, height = 400)
ggplot(melt(mod_m2), 
       aes(x=Var1, y=Var2, fill=factor(value))) + 
  geom_tile(colour="white") + 
  scale_fill_manual(values = c("#F1EEF6", "#D7B5D8", "#DF65B0", "#980043"), 
                    name = "Probability", 
                    labels = c("0 : 0.0004", "0.0004 : 0.0014", "0.0014 : 0.0052", "0.0052 : 0.0958")) +
  theme_grey(base_size = base_size) + 
  labs(x = "2010 Neighborhood Type", y = "2000 Neighborhood Type") +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(axis.ticks=element_blank(),
        axis.title.x = element_text(size = 18, face = "bold"),
        axis.title.y = element_text(size = 18, face = "bold"))
#dev.off()
```
